---
title: 'Problem Set 6  #1'
author: "Group 1"
date: "November 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

## 1.1
```{r 1.1.setup}
houseprice = read.table("hprice1.raw")
names(houseprice) = c("price","assess","bdrms","lotsize","sqrft","colonial","lprice","lassess","llotsize","lsqrft")
model = lm(price~sqrft+bdrms, data=houseprice)
model
```


#a.)
```{r 1.1.a}
head(houseprice)
summary(houseprice)
summary(subset(houseprice, select=c("price","sqrft","bdrms")))
```


#b.)
```{r 1.1.b}
Y = houseprice$price
X = cbind(rep(1,nrow(houseprice)),houseprice$sqrft,houseprice$bdrms)
```


The matrix notation is Y = XB + e, where the first five rows look like

$$\left(\begin{array}
{rrr}
300 \\
370 \\
191 \\
195 \\
373 \\
....
\end{array}\right)
\ = \left(\begin{array}
{rrr}
1 & 2438 &  4\\
1 & 2078 &  3\\
1 & 1374 &  3\\
1 & 1448 &  3\\
1 & 2514 &  4\\
...& ..... & ...
\end{array}\right)
\left(\begin{array}
{rrr}
Beta0 \\
Beta1 \\
Beta2 \\
\end{array}\right)
\ + \left(\begin{array}
{rrr}
epsilon1 \\
epsilon2 \\
epsilon3 \\
epsilon4 \\
epsilon5 \\
........
\end{array}\right)
$$

#c.)
```{r 1.1.c}
marg.inc.bdrm = model$coef[3]
```
Ceteris paribus, we expect house price to raise $`r round(marg.inc.bdrm*1000,2)` for every additional bedroom.


#d.)
```{r 1.1.d}
marg.inc.bdrm140 = model$coef[3] + 140*model$coef[2]
```
Ceteris paribus, we expect house price to raise $`r round(marg.inc.bdrm140*1000,2)` for an additional 140 sqft bedroom.


#e.)
```{r 1.1.e}
summary(model)
```
By the adjusted R^2, `r round(summary(model)$adj.r.squared*100,2)`% of variation in price is explained by sq footage and number of bedrooms.


#f.)
```{r 1.1.f}
yhats = predict(model)
```
The predicted selling price for the first house is $`r round(yhats[1]*1000,2)`.


#g.)
```{r 1.1.g}
resids = houseprice$price-yhats
#Ordinally, I would use model$residuals, but we are told to "calculate" it.
```
The residual for the first house using this model is `r resids[1]`. Therefore, the buyer underpaid for the house by $`r abs(round(resids[1]*1000,2))`.


#h.)
```{r 1.1.h}
```



## 1.2
#a.)
```{r 1.2.setup}
subs401k = read.table("401ksubs.raw")
names(subs401k) = c("e401k","inc","marr","male","age","fsize","nettfa","p401k","pira","incsq","agesq")
head(subs401k)
```


```{r 1.2.a}
clean.subs401k = subset(subs401k, fsize==1)
sing.per.house = nrow(clean.subs401k)
```
There are `r sing.per.house` single-person households in the data set.


```{r 1.2.b}
initial.model = lm(nettfa~inc+age, data=clean.subs401k)
initial.model
```
Ceteris paribus, we expect net total financial assets to increase \$`r round(initial.model$coef[2]*1000,2)` for every $1000 increase in annual income.

Ceteris paribus, we expect net total financial assets to increase \$`r round(initial.model$coef[3]*1000,2)` for every additional year older the respondent is.


```{r 1.2.c}
summary(initial.model)
```
